---
title: "analysis 3.12"
output: html_document


install.packages("ARDL")

"quantmod", "tsDyn", "vrtest", "nardl")
---

# Load packages
```{r}
library(dynlm)
library(zoo)
library(readxl)
library(tidyverse)
library(lmtest)
library(TTR)
library(sandwich)
library(car)
library(tseries)
library(urca)
library(ecm)
library(forecast)
library(knitr)
library(modelr)
library(astsa)
library(tsbox)
library(vars)
library(dLagM)
library(quantmod)
library(tsDyn)
library(vrtest)
library(ARDL)
library(nardl)
```

# Import data file
```{r}
cyprus <- read_excel("file.path")

cyprus <- cyprus_75_16
```

# Log/square variables
```{r}
library(dplyr)
cyprus <- cyprus %>% mutate(GDP2 = GDP*GDP)
cyprus <- cyprus %>% mutate(LCO2 = log(CO2))
cyprus <- cyprus %>% mutate(LGDP = log(GDP))
cyprus <- cyprus %>% mutate(LGDP2 = (log(GDP))*(log(GDP)))
cyprus <- cyprus %>% mutate(LFDI = log(FDI))
constant <- 1-min(cyprus$TRD)
cyprus <- cyprus %>% mutate(LTRD = log(TRD+constant))
cyprus <- cyprus %>% mutate(LENER = log(ENER))
cyprus <- cyprus %>% mutate(LRENEW = log(RENEW))
```

# Simple OLS and diagnostics
```{r}
fit <- lm(LCO2 ~ LGDP + LGDP2 + LFDI + LTRD + LENER, data = cyprus)
summary(fit)
durbinWatsonTest(fit)
dwtest(fit)
bgtest(fit)

residual <- resid(fit)
lag.plot(residual, 1, diag.col = "forest green", main = "Lag Scatter Plot")
lag1.plot(residual, 1)
cyprus <- add_residuals(cyprus, fit, var = "resid")
```

# Unit root test
```{r}
pacf(cyprus$LCO2)
adf.test(cyprus$LCO2, k=1)
pacf(cyprus$LGDP)
adf.test(cyprus$LGDP, k=1)
pacf(cyprus$LGDP2)
adf.test(cyprus$LGDP2, k=1)
pacf(cyprus$LFDI)
adf.test(cyprus$LFDI, k=1)
pacf(cyprus$LTRD)
adf.test(cyprus$TRD, k=1)
pacf(cyprus$LENER)
adf.test(cyprus$LENER, k=1)
```

## Constant vs. non-constant variance of independent variables?
```{r}
detach("package:dplyr", unload = TRUE)
kvec <- c(2, 5, 10)
cyprus.ts <- ts(cyprus, start = c(1975,1), end = c(2015,1), frequency = 1)
Lo.Mac(cyprus.ts[,"LCO2"], kvec)
Wald(cyprus.ts[,"LCO2"], kvec)
Lo.Mac(cyprus.ts[,"LGDP"], kvec)
Wald(cyprus.ts[,"LGDP"], kvec)
Lo.Mac(cyprus.ts[,"LGDP2"], kvec)
Wald(cyprus.ts[,"LGDP2"], kvec)
Lo.Mac(cyprus.ts[,"LFDI"], kvec)
Wald(cyprus.ts[,"LFDI"], kvec)
Lo.Mac(cyprus.ts[,"LTRD"], kvec)
Wald(cyprus.ts[,"LTRD"], kvec)
Lo.Mac(cyprus.ts[,"LENER"], kvec)
Wald(cyprus.ts[,"LENER"], kvec)
library(dplyr)
```

# Calculating ARDL

## Calculate optimal lag order
```{r}
model1 <- auto_ardl(LCO2 ~ LGDP + LGDP2 + LFDI + LTRD + LENER, data = cyprus, max_order = 3, selection = "BIC")
model1$top_orders
```

## Simple (unrestricted) ARDL model
```{r}
ardl <- ardl(LCO2 ~ LGDP + LGDP2 + LFDI + LTRD + LENER, data = cyprus, order = c(1,0,1,0,0,1))
summary(ardl)
```

## t-bound test
```{r}
ardl <- ardl(LCO2 ~ LGDP + LGDP2 + LFDI + LTRD + LENER, data = cyprus, order = c(1,0,1,0,0,1))
summary(ardl)
tbounds <- bounds_t_test(ardl, case = 3, alpha = 0.01)
tbounds
tbounds$tab
```
## f-bound test
```{r}
ardl <- ardl(LCO2 ~ LGDP + LGDP2 + LFDI + LTRD + LENER, data = cyprus, order = c(1,0,1,0,0,1))
summary(ardl)
fbounds <- bounds_f_test(ardl, case = 3)
fbounds
```
## "Cointegration test for cointegration equation" & unrestricted ECM

```{r}
ardl <- ardl(LCO2 ~ LGDP + LGDP2 + LFDI + LTRD + LENER, data = cyprus, order = c(1,0,1,0,0,1))
ce2_ardl <- coint_eq(ardl, case = 3)
uecm <- uecm(ardl)
summary(uecm)
ce2_uecm <- coint_eq(uecm, case = 3)
```
## Restricted ECM
```{r
recm <- recm(uecm, case = 3)
summary(recm)
```
## Check if identical
```{r}
identical(ce2_ardl, ce2_uecm)
```
## Get long-run estimates
```{r}
mult_ardl <- multipliers(ardl)
mult_ardl
```

## Plot
```{
output <- cbind.zoo(LRM = cyprus$LCO2, ce2_recm)
plot(output, lty = c(2, 1), lwd = 2, xlab = "Year", main = "CO2")
```
#Stability tests
```{r}
ardl1 <- ardlBound(LCO2 ~ LGDP + LGDP2 + LFDI + LTRD + LENER, data = cyprus, case = 3, p = 1, max.q = 1, ECM = FALSE, stability = TRUE)
```

# Granger test
```{r}
grangertest(LCO2 ~ LGDP, order = 1, data = cyprus)
grangertest(LGDP ~ LCO2, order = 1, data = cyprus)
grangertest(LCO2 ~ LGDP2, order = 1, data = cyprus)
grangertest(LGDP2 ~ LCO2, order = 1, data = cyprus)
grangertest(LCO2 ~ LFDI, order = 1, data = cyprus)
grangertest(LFDI ~ LCO2, order = 1, data = cyprus)
grangertest(LCO2 ~ LTRD, order = 1, data = cyprus)
grangertest(LTRD ~ LCO2, order = 1, data = cyprus)
grangertest(LCO2 ~ LENER, order = 1, data = cyprus)
grangertest(LENER ~ LCO2, order = 1, data = cyprus)
```

